---
title: "pFDA-GIAB 2nd Challenge Results"
date: '`r Sys.Date()`'
output: 
    bookdown::html_document2:
        toc: true
        toc_float: true
        df_print: paged
        code_folding: hide
---

```{r message = FALSE}
library(tidyverse)
library(readxl)
library(DT)
library(sessioninfo)
library(here)
library(ggrepel)
source(here("scripts", "annotation-phredticks.R"))
```

# Summary
Summarize challenge results for identification of winners and reporting overall performance.
Winners identified for the four technologies and three stratifications based on the geometric mean of the F1 metric scores for the HG003 and HG004 submissions. 

# Loading and tidying data

## Participant metadata
```{r}
participant_df <- here("data-raw", "anonymized_metadata_table.tsv") %>% 
    read_tsv() 
coauthor_subs_df <- participant_df %>% 
    filter(methods_provided == "Y")
```


```{r}
participant_df %>% 
DT::datatable(caption = "Truth challenge submission metadata.", rownames = FALSE)
```

## Benchmarking Results
```{r}
results_df <- read_tsv(here("data","anonymized_challenge_results_v5.txt")) %>% 
    ## Only including submissions from co-authors
    semi_join(coauthor_subs_df) %>% 
    mutate(challenge_cat = factor(
        challenge_cat,
        levels = c("Difficult-to-Map Regions",
                   "All Benchmark Regions",
                   "MHC")
        )
    ) %>%
    ## Adding phred version of metrics
    mutate(phred_F1 = -10*log(1-METRIC.F1_Score)/log(10),
           phred_Precision = -10*log(1-METRIC.Precision)/log(10),
           phred_Recall = -10*log(1-METRIC.Recall)/log(10)) %>% 
    filter(Type == "SNVandINDEL")
```


```{r}
results_df %>% 
    select(`Participant(s)`, Submission_Name, Technology, challenge_cat) %>%
DT::datatable(caption = "Challenge Results.", rownames = FALSE)
```

## Assigning Winners
There are 13 rows in the table, rather than the expected 12 - 4 technologies X 3 stratification, because 2 of Hanying Feng's submissions had the same score. 
```{r}
# winners_tbl <- results_df %>% 
#     group_by(challenge_cat, Technology) %>% 
#     top_n(1, METRIC.F1_Score)
# 
# tied_winner_df <- results_df %>% 
#     filter(`Participant(s)` == "Mohammad Sahraeian",
#            Technology == "MULTI",
#            challenge_cat == "All Benchmark Regions",
#            Submission_Name == "RN-Illumina-PacBio-ONT")
# 
# winners_tbl <- bind_rows(winners_tbl, tied_winner_df) %>% 
#     select(`Participant(s)`, Technology, challenge_cat, METRIC.F1_Score, Submission_Name) %>% 
#     mutate(METRIC.F1_Score = round(METRIC.F1_Score, 4)) %>% 
#     arrange(-METRIC.F1_Score) %>% 
#     rename(F1 = "METRIC.F1_Score")
#
## Saving table with challenge winners to file.
# winners_tbl %>% 
#     mutate(Category = paste(challenge_cat, Technology, sep = " - ")) %>% 
#     ungroup() %>% 
#     select(Category, `Participant(s)`, Submission_Name, F1) %>% 
# write_tsv(here("data","challenge_winners.tsv"))
```

Loading saved winners table
```{r}
winners_tbl <- read_tsv(here("data","challenge_winners.tsv"))
```

```{r}
winners_tbl %>% DT::datatable(caption = "Challenge Winners", rownames = FALSE)
```

Checking for honorable mentions: 

- Illumina: clear winner for all categories
- PacBio: Three of Andrew Carroll's submissions had similar performance (differed in fourth decimal place)
- ONT: Kishwar lowmap similar for ont (0.964633) and ont_9to6 (0.965146) 
- Multi: All - Hanying Ill+PacBio (0.998995), Hanying Ill+PacBio+ONT (0.998978), and Mohammad Sahreaian (0.998978) were very close, MHC - Hanying Ill+PacBio+ONT two models were tied (0.997869585420914)

```{r}
results_df %>% 
    group_by(challenge_cat, Technology) %>% 
    mutate(METRIC.F1_Score = round(METRIC.F1_Score, 3)) %>% 
    top_n(3, METRIC.F1_Score) %>% 
    select(`Participant(s)`, Technology, challenge_cat, METRIC.F1_Score, Submission_Name) %>% 
    arrange(-METRIC.F1_Score) %>% 
    rename(`Genomic Region` = "challenge_cat", F1 = "METRIC.F1_Score") %>% 
    arrange(Technology, `Genomic Region`, -F1) %>% 
DT::datatable(caption = "Challenge winners, one winner selected per technology and stratification; All - GIAB 4.2 benchmark regions, MHC - major histocompatibility region, and LowMap - combined 100 bp low mappability and segmental duplications.", rownames = FALSE)
```
## Reformatting Winners Table
```{r}
## Adding winners to results df for plot annotation
results_annotated_df <- winners_tbl %>%
    separate(Category, c("challenge_cat", "Technology"), sep = " - ") %>%
    mutate(
        Submission_Name = if_else(
            str_detect(Submission_Name, "DRAGEN"),
            "DRAGEN Experimental Extension into Difficult Regions",
            Submission_Name
        ),
        challenge_cat = factor(
            challenge_cat,
            levels = c("Difficult-to-Map Regions",
                       "All Benchmark Regions",
                       "MHC")
        )
    ) %>%
    add_column(winners = TRUE) %>%
    right_join(results_df)
```


```{r}
(ms_winners_tbl <- results_annotated_df %>% 
     mutate(winners = if_else(Technology == "MULTI" & 
                                  Submission_Name == "DeepVariant Hybrid PacBio-Illumina" & 
                                  challenge_cat == "All Benchmark Regions", TRUE, winners)) %>% 
    filter(winners == TRUE) %>% 
     mutate(Recall = round(METRIC.Recall, 4),
            Precision = round(METRIC.Precision, 4)) %>% 
     select(challenge_cat, Technology, 
            Submitter, Submission_Name, `Participant(s)`,
            F1, Recall, Precision) %>% 
    mutate(Submission_Name = if_else(str_detect(Submission_Name, "DRAGEN"),
                                     "DRAGEN Experimental Extension into Difficult Regions",
                                     Submission_Name)) %>% 
    left_join(results_df) %>% 
    select(Technology, challenge_cat, `Participant(s)`, F1, Recall, Precision) %>% 
         mutate(Technology = factor(Technology, 
                               levels = c("MULTI", "ILLUMINA","PACBIO","ONT"))
           ) %>%
    arrange(Technology, challenge_cat) %>% 
    rename(`Genomic Region` = challenge_cat,
           Participant = `Participant(s)`) %>% 
    distinct())

# write_csv(ms_winners_tbl, here("figures","ms_winners_tbl.csv"))
```
```{r}
results_annotated_df %>% filter(challenge_cat == "All Benchmark Regions", Technology == "MULTI")
```

Getting Ranks for annotating Table 2
```{r}
results_annotated_df %>% 
     ## Adding F1 ranks
    group_by(challenge_cat, Technology) %>% 
    mutate(F1_rank = rank(-round(METRIC.F1_Score,3),ties.method = "min")) %>% 
     mutate(winners = if_else(Technology == "MULTI" & 
                                  Submission_Name == "DeepVariant Hybrid PacBio-Illumina" & 
                                  challenge_cat == "All Benchmark Regions", TRUE, winners)) %>% 
    mutate(Submission_Name = if_else(str_detect(Submission_Name, "DRAGEN"),
                                     "DRAGEN Experimental Extension into Difficult Regions",
                                     Submission_Name)) %>% 
    mutate(F1_rank_cat = str_sub(challenge_cat, 1,4)) %>% 
    ungroup() %>% 
    select(Technology, `Participant(s)`, F1_rank, F1_rank_cat, Submission_Name)  %>% 
    arrange(Technology, F1_rank_cat, F1_rank)
```


# Overall Results
```{r}
f1_dist_plot
```


## F1
```{r }
## Metric distributions by stratification ########
f1_dist_plot <- results_df %>%
    ggplot() +
    geom_histogram(aes(x = phred_F1), alpha = 0.5) +
    geom_histogram(
        aes(x = phred_F1, fill = Technology),
        size = 0.25,
        color = "grey20",
        position = "dodge",
        na.rm = TRUE
    ) +
    facet_wrap( ~ challenge_cat, ncol = 1) +
    scale_fill_brewer(type = "qual", palette = 2) +
    scale_x_continuous(breaks = c(0, 10, 20, 30),
                       labels = c("0", "90", "99", "99.9")) +
    annotation_phredticks(sides = "b",
                          outside = TRUE,
                          scaled = FALSE) +
    scale_y_continuous(breaks = scales::pretty_breaks()) +
    coord_cartesian(clip = "off") +
    theme_bw() +
    theme(axis.text.x = element_text(vjust = -1.5)) +
    labs(x = "F1 %", y = "Count")


## Submission Performance #########################
f1_trend_plot <- results_annotated_df %>% 
    ggplot() +
    geom_line(
        aes(
            x = as.integer(challenge_cat),
            y = phred_F1,
            group = Submission_Name,
            color = Technology
        ),
        size = 2,
        alpha = 0.25
    ) +
    geom_point(
        aes(
            x = as.integer(challenge_cat),
            y = phred_F1,
            fill = Technology
        ),
        shape = 21,
        alpha = 0.5
    ) +
    geom_point(
        data = filter(results_annotated_df, winners),
        aes(
            x = as.integer(challenge_cat),
            y = phred_F1,
            fill = Technology,
        ),
        size = 5,
        shape = 23,
        show.legend = FALSE
    ) +
    geom_point(
        data = filter(results_annotated_df, winners),
        aes(
            x = as.integer(challenge_cat),
            y = phred_F1,
            fill = Technology,
        ),
        size = 3,
        shape = "W",
        show.legend = FALSE
    ) +
    geom_text_repel(
            data = filter(results_annotated_df, winners),
            point.padding = 0.25,
            aes(x = as.integer(challenge_cat),
                y = phred_F1,
                label = `Participant(s)`),
            na.rm = TRUE,
            min.segment.length = 0.2,
            size = 3
        ) +
    scale_x_continuous(breaks = 1:3,
                       labels = c("Difficult-to-Map\nRegions",
                                  "All Benchmark\nRegions",
                                  "MHC")) +
    scale_y_continuous(breaks = c(0, 10, 20, 30),
                       labels=c("0","90","99","99.9")) +
    annotation_phredticks(sides = "lr", scaled = FALSE) +
    scale_fill_brewer(type = "qual", palette = 2) +
    scale_color_brewer(type = "qual", palette = 2) +
    theme_bw() +
    theme(legend.position = "right") +
    labs(x = "Genomic Regions", y = "F1 %")
```

Saving plot objects for use in combined plot in `challenge-overview.Rmd`
```{r}
saveRDS(f1_dist_plot, "f1_dist_plot.RDS")
saveRDS(f1_trend_plot, "f1_trend_plot.RDS")
```


__Combined Plot__
```{r}
(f1_combined_plt <- ggpubr::ggarrange(f1_dist_plot, f1_trend_plot, nrow = 2,
                  common.legend = TRUE, 
                  legend = "bottom", labels = "AUTO"))
```


```{r}
ggsave(filename = here("figures","challenge_results_f1.pdf"), 
       f1_combined_plt,width = 8, height = 8
       )

ggsave(filename = here("figures","challenge_results_f1.png"), 
       f1_combined_plt,width = 8, height = 8
       )
```


## Precision
```{r }
## Metric distributions by stratification ########
precision_dist_plot <- results_df %>%
    ggplot() +
    geom_histogram(aes(x = phred_Precision), alpha = 0.5) +
    geom_histogram(
        aes(x = phred_Precision, fill = Technology),
        color = "grey20",
        position = "dodge",
        na.rm = TRUE
    ) +
    facet_wrap( ~ challenge_cat, ncol = 1) +
    scale_fill_brewer(type = "qual", palette = 2) +
    scale_x_continuous(breaks = c(0, 10, 20, 30),
                       labels = c("0", "90", "99", "99.9")) +
    annotation_phredticks(sides = "b",
                          outside = TRUE,
                          scaled = FALSE) +
    scale_y_continuous(breaks = scales::pretty_breaks()) +
    coord_cartesian(clip = "off") +
    theme_bw() +
    theme(axis.text.x = element_text(vjust = -1.5)) +
    labs(x = "Precision %", y = "Count")


precision_trend_plot <- results_annotated_df %>% 
    ggplot() +
    geom_line(
        aes(
            x = as.integer(challenge_cat),
            y = phred_Precision,
            group = Submission_Name,
            color = Technology
        ),
        size = 2,
        alpha = 0.25
    ) +
    geom_point(
        aes(
            x = as.integer(challenge_cat),
            y = phred_Precision,
            fill = Technology
        ),
        shape = 21,
        alpha = 0.5
    ) +
    geom_point(
        data = filter(results_annotated_df, winners),
        aes(
            x = as.integer(challenge_cat),
            y = phred_Precision,
            fill = Technology,
        ),
        size = 5,
        shape = 23,
        show.legend = FALSE
    ) +
    geom_point(
        data = filter(results_annotated_df, winners),
        aes(
            x = as.integer(challenge_cat),
            y = phred_Precision,
            fill = Technology,
        ),
        size = 3,
        shape = "W",
        show.legend = FALSE
    ) +
    geom_text_repel(
            data = filter(results_annotated_df, winners),
            point.padding = 0.25,
            aes(x = as.integer(challenge_cat),
                y = phred_Precision,
                label = `Participant(s)`),
            na.rm = TRUE,
            min.segment.length = 0.2,
            size = 3
        ) +
    scale_x_continuous(breaks = 1:3,
                       labels = c("Difficult-to-Map\nRegions",
                                  "All Benchmark\nRegions",
                                  "MHC")) +
    scale_y_continuous(breaks = c(0, 10, 20, 30),
                       labels=c("0","90","99","99.9")) +
    annotation_phredticks(sides = "lr", scaled = FALSE) +
    scale_fill_brewer(type = "qual", palette = 2) +
    scale_color_brewer(type = "qual", palette = 2) +
    theme_bw() +
    theme(legend.position = "right") +
    labs(x = "Genomic Regions", y = "Precision %")
```


__Combined Plot__
```{r}
(
    precision_combined_plt <- ggpubr::ggarrange(
        precision_dist_plot,
        precision_trend_plot,
        nrow = 2,
        common.legend = TRUE,
        legend = "bottom",
        labels = "AUTO"
    )
)
```

```{r}
ggsave(filename = here("figures","challenge_results_precision.pdf"), 
       precision_combined_plt, width = 8, height = 8
       )

ggsave(filename = here("figures","challenge_results_precision.png"), 
       precision_combined_plt, width = 8, height = 8
       )
```

## Recall
```{r }
## Metric distributions by stratification ########
recall_dist_plot <- results_df %>%
    ggplot() +
    geom_histogram(aes(x = phred_Recall), alpha = 0.5) +
    geom_histogram(
        aes(x = phred_Recall, fill = Technology),
        color = "grey20",
        position = "dodge",
        na.rm = TRUE
    ) +
    facet_wrap( ~ challenge_cat, ncol = 1) +
    scale_fill_brewer(type = "qual", palette = 2) +
    scale_x_continuous(breaks = c(0, 10, 20, 30),
                       labels = c("0", "90", "99", "99.9")) +
    annotation_phredticks(sides = "b",
                          outside = TRUE,
                          scaled = FALSE) +
    scale_y_continuous(breaks = scales::pretty_breaks()) +
    coord_cartesian(clip = "off") +
    theme_bw() +
    theme(axis.text.x = element_text(vjust = -1.5)) +
    labs(x = "Recall %", y = "Count")

recall_trend_plot <- results_annotated_df %>% 
    ggplot() +
    geom_line(
        aes(
            x = as.integer(challenge_cat),
            y = phred_Recall,
            group = Submission_Name,
            color = Technology
        ),
        size = 2,
        alpha = 0.25
    ) +
    geom_point(
        aes(
            x = as.integer(challenge_cat),
            y = phred_Recall,
            fill = Technology
        ),
        shape = 21,
        alpha = 0.5
    ) +
    geom_point(
        data = filter(results_annotated_df, winners),
        aes(
            x = as.integer(challenge_cat),
            y = phred_Recall,
            fill = Technology,
        ),
        size = 5,
        shape = 23,
        show.legend = FALSE
    ) +
    geom_point(
        data = filter(results_annotated_df, winners),
        aes(
            x = as.integer(challenge_cat),
            y = phred_Recall,
            fill = Technology,
        ),
        size = 3,
        shape = "W",
        show.legend = FALSE
    ) +
    geom_text_repel(
            data = filter(results_annotated_df, winners),
            point.padding = 0.25,
            aes(x = as.integer(challenge_cat),
                y = phred_Recall,
                label = `Participant(s)`),
            na.rm = TRUE,
            min.segment.length = 0.2,
            size = 3
        ) +
    scale_x_continuous(breaks = 1:3,
                       labels = c("Difficult-to-Map\nRegions",
                                  "All Benchmark\nRegions",
                                  "MHC")) +
    scale_y_continuous(breaks = c(0, 10, 20, 30),
                   labels=c("0","90","99","99.9")) +
    annotation_phredticks(sides = "lr", scaled = FALSE) +
    scale_fill_brewer(type = "qual", palette = 2) +
    scale_color_brewer(type = "qual", palette = 2) +
    theme_bw() +
    theme(legend.position = "right") +
    labs(x = "Genomic Regions", y = "Recall %")
```

__Combined Plot__
```{r}
(
    recall_combined_plt <- ggpubr::ggarrange(
        recall_dist_plot,
        recall_trend_plot,
        nrow = 2,
        common.legend = TRUE,
        legend = "bottom",
        labels = "AUTO"
    )
)
```

```{r}
ggsave(filename = here("figures","challenge_results_recall.pdf"), 
       recall_combined_plt, width = 8, height = 8
       )

ggsave(filename = here("figures","challenge_results_recall.png"), 
       recall_combined_plt, width = 8, height = 8
       )
```

# Session Information
## System Information
```{r}
sessioninfo::platform_info()
```


## Package Versions
```{r}
sessioninfo::package_info() %>% 
    filter(attached = TRUE) %>% 
    select(package, loadedversion, date, source) %>%
    knitr::kable(booktabs = TRUE, row.names = FALSE)
```

