---
title: "pFDA-GIAB 2nd Challenge Results"
date: '`r Sys.Date()`'
output: 
    bookdown::html_document2:
        toc: true
        toc_float: true
        df_print: paged
        code_folding: hide
---


```{r message = FALSE}
library(tidyverse)
library(here)
```

# Objective
Identification of difficult stratification for identification of challenge winners.

## Proposed Analyses

- __Identification of difficult stratification__: Compare the rankings for LowMap250bp.SNVINDEL.F1, LowMap100bpSegDup.SNVINDEL.F1, and SegDup.SNVINDEL.F1, and decide which to use - hopefully there aren't major differences and there's a clear winner in each tech category. _See Analysis section below.


# Loading and tidying data

Summarized results
```{r}
results_df <- here("data-raw", "FinalTruthExtracted.xlsx") %>% 
    readxl::read_xlsx() %>% 
    ## Converting to a long data frame
    select(-Extra_Details, -Participant_File) %>% 
    gather("metric_cat","value", 
           -Name, -Submission_Name, -HG_Num, -Technology, -Multi_Details) %>% 
    
```

## Tidying results for data analysis

Tidying metric names for parsing

```{r}
results_df <- results_df %>% 
    mutate(metric_cat = str_replace(metric_cat, "AllSNV", "All_SNV"),
           metric_cat = str_replace(metric_cat, "SNV_INDEL", "SNV-INDEL"),
           metric_cat = str_replace(metric_cat, "LowMap_", "LowMap-"),
           metric_cat = str_replace(metric_cat, "bp_SegDup", "bp-SegDup")) %>% 
    separate(metric_cat, c("strat","var_type","metric"), sep = "_", remove = FALSE)
```


```{r eval = FALSE}
## Checking metric category parsing
results_df %>% select(metric_cat, strat, var_type, metric) %>% 
    distinct() %>% 
    View()
```
## Defining Challenge categories
```{r}
results_df <- results_df %>% 
    mutate(challenge_cat = case_when(strat == "All" ~ "All",
                                     strat == "MHC" ~ "MHC",
                                     TRUE ~ "Difficult"))
```

```{r eval = FALSE}
## Checking challenge categories
results_df %>% 
    select(metric_cat, strat, var_type, metric, challenge_cat) %>% 
    distinct() %>% 
    View()
```

## Calculating mean metric value
```{r}
results_df <- results_df %>% 
    filter(HG_Num != "002") %>% 
    group_by(Name, Technology, Multi_Details, Submission_Name, 
             metric_cat, strat, var_type, metric, challenge_cat) %>% 
    summarise(value_mean = exp(mean(log(value))))
```

# Analyses


## Identification of Difficult Stratification

Comparing variant callset (submission) ranks across the three difficult stratifications (lowMap 100 bp + seg dups, lowMap 250 bp, and segdups), to determine which stratification to use for the winner assignment. 

Three analyses for echo technology category (Illumina, PacBio, ONT, and multiple).  

1. Comparison of submission ranks (geometric mean for HG003 and HG004 F1 metric for combined SNP and INDELS) across the three difficult stratifications. 
2. Comparison of submission F1 metric values across difficult stratifications.
3. Correlation analysis - indicates extent to which challenge rankings change based on which stratification is used. 

```{r}
## Formatting data frame for figures
top_diff_results_df <- results_df %>% 
    filter(challenge_cat == "Difficult", var_type == "SNV-INDEL") %>% 
    group_by(Technology, var_type, metric, strat) %>% 
    arrange(-value_mean) %>% 
    mutate(rank = 1:n()) %>% 
    ## formatting for figure
    mutate(Multi_Details = if_else(is.na(Multi_Details), "", Multi_Details), 
           sub_id = str_c(Name, Technology, Multi_Details, Submission_Name)) %>% 
    mutate(strat = factor(strat, levels = c("LowMap-250bp","LowMap-100bp-SegDup","SegDup")))

## Stratification plot
diff_strat_comp_plot <- function(df){
        ggplot(df) + 
        geom_line(aes(x = as.integer(strat), y = rank, group = sub_id)) + 
        geom_point(aes(x = as.integer(strat), y = rank)) + 
        facet_grid(Technology~., scales = "free_y") + 
        scale_x_continuous(breaks = 1:3, 
                           labels = c("LowMap-250bp","LowMap-100bp-SegDup","SegDup")) +
        theme_bw() + 
        theme(legend.position = "bottom") + 
        labs(x = "Stratification", y = "Submission Rank")
}

diff_strat_comp_value_plot <- function(df){
        ggplot(df) + 
        geom_line(aes(x = as.integer(strat), y = 1- value_mean, group = sub_id)) + 
        geom_point(aes(x = as.integer(strat), y = 1- value_mean)) + 
        facet_grid(Technology~., scales = "free_y") + 
        scale_x_continuous(breaks = 1:3, 
                           labels = c("LowMap-250bp","LowMap-100bp-SegDup","SegDup")) +
        scale_y_log10() +
        theme_bw() + 
        theme(legend.position = "bottom") + 
        labs(x = "Stratification", y = "log10(1-F1)")
}

## Stratification Correlations
get_strat_corrs <- function(df){
    strat_mat <- df %>% ungroup() %>%
        select(sub_id, strat, rank) %>%
        spread(strat, rank) %>%
        select(-sub_id) %>%
        as.matrix()
    
    cor_mat <- cor(strat_mat)
    cor_mat
    cor_means <- colMeans(cor_mat)
    cor_means_df <- t(cor_means) %>% 
        as.data.frame() %>% 
        add_column(strat = "Mean",.before = 1)
    
    ## Combining into a single data frame
    cor_mat %>% 
        as.data.frame() %>% 
        rownames_to_column(var = "strat") %>% 
        bind_rows(cor_means_df)
}

```

### Illumina
```{r}
tech_df <- filter(top_diff_results_df,Technology == "ILLUMINA") 
tech_df %>% diff_strat_comp_plot()
```

```{r}
diff_strat_comp_value_plot(tech_df)
```

```{r}
tech_df %>% get_strat_corrs()
```


### PacBio
```{r}
tech_df <- filter(top_diff_results_df,Technology == "PACBIO") 
tech_df %>% diff_strat_comp_plot()
```
```{r}
diff_strat_comp_value_plot(tech_df)
```

```{r}
tech_df %>% get_strat_corrs()
```

### ONT
```{r}
tech_df <- filter(top_diff_results_df,Technology == "ONT") 
tech_df %>% diff_strat_comp_plot()
```
The two top caller results are overlapping.
```{r}
diff_strat_comp_value_plot(tech_df)
```

```{r}
tech_df %>% get_strat_corrs()
```

### Multiple Technologies
```{r}
tech_df <- filter(top_diff_results_df,Technology == "MULTI") 
tech_df %>% diff_strat_comp_plot()
```

```{r}
diff_strat_comp_value_plot(tech_df)
```

```{r}
tech_df %>% get_strat_corrs()
```




# Session Information
## System Information
```{r}
sessioninfo::platform_info()
```


## Package Versions
```{r}
sessioninfo::package_info() %>% 
    filter(attached = TRUE) %>% 
    select(package, loadedversion, date, source) %>%
    knitr::kable(booktabs = TRUE, row.names = FALSE)
```

