---
title: "Performance Summary for MHC"
date: '`r Sys.Date()`'
output: 
    bookdown::html_document2:
        toc: true
        toc_float: true
        df_print: paged
        code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(ggrepel)
library(ggpubr)
source(here("scripts", "annotation-phredticks.R"))
```

# Objective
Summarize submission performance for MHC region relative to other challenge categories and identification of variant callers that perform better in MHC regardless of overall performance. 


# Loading and Munging Data 
```{r}
coauthor_subs_df <- here("data-raw", "anonymized_metadata_table.tsv") %>% 
    read_tsv() %>% 
    filter(methods_provided == "Y") 

results_df <- read_tsv(here("data","anonymized_challenge_results_v5.txt")) %>% 
    ## Only including submissions from co-authors
    semi_join(coauthor_subs_df) %>% 
    mutate(challenge_cat = factor(
        challenge_cat,
        levels = c("Difficult-to-Map Regions",
                   "All Benchmark Regions",
                   "MHC")
        )
    ) %>%
    ## Adding PhredF1
    mutate(phred_F1 = -10*log(1-METRIC.F1_Score)/log(10)) 
```



```{r}
results_subset <- results_df %>% 
    select(Technology, Submission_Name, Type, 
           challenge_cat, METRIC.F1_Score, phred_F1)
    
results_MHC <- results_subset %>% 
    filter(challenge_cat == "MHC") %>% 
    rename(mhc_F1 = METRIC.F1_Score,
           mhc_phred_F1 = phred_F1) %>% 
    select(-challenge_cat)

results_comp <- results_subset %>% 
    filter(challenge_cat != "MHC") %>% 
    left_join(results_MHC) %>% 
    mutate(f1_diff = mhc_phred_F1 - phred_F1) %>% 
    group_by(Type, challenge_cat) %>% 
    arrange(-f1_diff) %>% 
    mutate(rank = 1:n(), 
           labeled_subs = if_else(rank < 2 & round(f1_diff,0) >= 0, Submission_Name, "")) %>% 
    mutate(labeled_subs = case_when(labeled_subs == "ONT NanoCaller Clair Medaka Ensemble" ~ "ONT Ensemble",
                                    labeled_subs == "Seven Bridges GRAF - Illumina" ~ "SevenBridges",
                                    labeled_subs == "PEPPER-DeepVariant (ONT_9to6)" ~ "PEPPER-DV",
                                    labeled_subs == "Combination of Illumina, PacBio HIFI, and Oxford Nanopore submission Model2" ~ "Combo Model 2", 
                                    labeled_subs == "Illumina-Pacbio-asmMHC v2(Genetalks)" ~ "asmMHC",
                                    labeled_subs == "Margin-PEPPER-DeepVariant (PacBIO)" ~ "PacBio Ensemble",
                                    labeled_subs == "Deepvariant_Clair_v1" ~ "DV+Clair",
                                    labeled_subs == "PacBio NanoCaller" ~ "NanoCaller",
                                    labeled_subs == "Pacbio-mm2SNP-bwaINDEL-asmMHC(Genetalks)" ~ "asmMHC",
                                    TRUE ~ labeled_subs)
           )
    
```

```{r}
results_comp %>% 
    filter(labeled_subs != "") %>% 
    arrange(Submission_Name)
```


```{r}
(faceted_plt <- results_comp %>% 
     mutate(alpha = if_else(labeled_subs == "", 0.25, 1)) %>% 
     mutate(Type = if_else(Type == "SNP", "SNV", Type)) %>% 
     ggplot(aes(x = phred_F1, y = mhc_phred_F1)) +
    geom_abline(aes(intercept = 0, slope = 1), color = "grey60") + 
    geom_point(aes(fill = Technology, 
                   label = Submission_Name, 
                   alpha = alpha), shape = 21) + 
    geom_text_repel(aes(label = labeled_subs), 
                  na.rm = TRUE, 
                  max.overlaps = 50,
                  min.segment.length = 0.2,
                  point.padding = 0.25) +
    scale_x_continuous(breaks = c(0, 10, 20, 30),
                   labels=c("0","90","99","99.9")) +
    scale_y_continuous(breaks = c(0, 10, 20, 30),
                   labels=c("0","90","99","99.9")) +
    facet_grid(Type ~ challenge_cat) + 
    theme_bw() + 
    theme(legend.position = "bottom") +
     scale_alpha_continuous(range = c(0.35,1)) +
    scale_fill_brewer(type = "qual", palette = 2) +
    annotation_phredticks(sides = "bl",
                          scaled = FALSE) +
    guides(alpha = "none") + 
    labs(x = "F1%", y = "MHC F1%"))
```

```{r}
ggsave(here("figures","mhc.pdf"), faceted_plt, width = 4.5, height = 6.75)
ggsave(here("figures","mhc.png"), faceted_plt, 
       width = 4.5, height = 6.75, dpi = "retina")
```

__Key points__  
- Illumina Graph based approach performs especially well in MHC. Add text about unique aspects of this graph based approach. 
- Submissions that performed especially poorly - was the MHC excluded or not called?
- NanoCaller Medaka ensemble performed well on MHC particularly for SNPs. Only method that performed as well in MHC as all benchmarking regions for SNPs.

```{r}
(graph_plt <- results_comp %>% 
     mutate(alpha = if_else(labeled_subs == "", 0.25, 1)) %>% 
     mutate(Type = if_else(Type == "SNP", "SNV", Type)) %>% 
          filter(Type == "SNVandINDEL") %>% 
    ggplot(aes(x = phred_F1, y = mhc_phred_F1)) + 
    geom_abline(aes(intercept = 0, slope = 1), color = "grey60") + 
    geom_point(aes(fill = Technology, 
                   label = Submission_Name, 
                   alpha = alpha), shape = 21) + 
    geom_text_repel(aes(label = labeled_subs), 
                  na.rm = TRUE, 
                  min.segment.length = 0.2,
                  point.padding = 0.25) +
    scale_x_continuous(breaks = c(0, 10, 20, 30), 
                   labels=c("0","90","99","99.9")) + 
    scale_y_continuous(breaks = c(0, 10, 20, 30), 
                   labels=c("0","90","99","99.9")) + 
    facet_grid(challenge_cat ~ Type) + 
    theme_bw() + 
    theme(legend.position = "bottom") +
     scale_alpha_continuous(range = c(0.35,1)) +
    scale_fill_brewer(type = "qual", palette = 2) +
    guides(alpha = FALSE) + 
    labs(x = "F1% (log)", y = "MHC F1%"))
```

```{r}
ggsave(here("figures","mhc_SNVandINDEL-only.pdf"), graph_plt, width =5, height = 4.5)
```


# Exploratory Plots
```{r include=FALSE}
results_df %>% 
    ggplot() + geom_point(aes(x = Submission_Name, y = phred_F1, shape = Technology, color = Type)) + 
    facet_grid(.~challenge_cat)
```

```{r include=FALSE}
ggdotchart(filter(results_df, challenge_cat == "MHC"),
           x = "Submission_Name", y = "phred_F1",
           color = "Type",                                # Color by groups
           # palette = c("#00AFBB", "#E7B800", "#FC4E07"), # Custom color palette
           sorting = "descending",                       # Sort value in descending order
           add = "segments",                             # Add segments from y = 0 to dots
           rotate = FALSE,                                # Rotate vertically
           group = "Technology",                                # Order by groups
           shape = "Technology",
           # dot.size = 2,                                 # Large dot size
           font.label = list(color = "grey20", size = 8, 
                             vjust = 0.5)              # Adjust label parameters
           ) + 
    # scale_y_continuous(trans='log2') + 
    # theme_bw() +
        theme(axis.text.x = element_blank(),
          legend.position = "bottom", legend.box = "bottom") +
    scale_color_brewer(type = "qual", palette = 2) + 
    labs(x = "Submissions", y = "F1% (log)",
         color = "Variant Caller Category")
```

## Scatter Plots
```{r include=FALSE}
all_v_MHC <- results_df %>% 
    select(`Participant(s)`, Technology, Multi_Details, Submission_Name, 
           Type, challenge_cat, phred_F1) %>% 
    pivot_wider(names_from = "challenge_cat", 
                values_from = "phred_F1") %>% 
    ggplot() + 
    geom_abline(aes(intercept = 0, slope = 1)) + 
    geom_point(aes(x = `All Benchmark Regions`, y = MHC, 
                   fill = Technology, label = Submission_Name), 
               shape = 21) + 
    geom_text_repel(point.padding = 0.25, aes(x = V3.3.2_unique, y = V4.1_unique, label = Subset), 
                  size = 3) +

    scale_x_continuous(breaks = c(0, 10, 20, 30), 
                   labels=c("0","90","99","99.9")) + 
    scale_y_continuous(breaks = c(0, 10, 20, 30), 
                   labels=c("0","90","99","99.9")) + 
    facet_wrap(~Type) + 
    theme_bw()  + 
    theme(legend.position = "bottom") +
    scale_fill_brewer(type = "qual", palette = 2)
```

### Combined Plots
```{r include=FALSE}
# (combined_plt <- ggarrange(all_v_MHC, diff_v_MHC, 
#                           nrow = 2, 
#                            labels = "AUTO", 
#                             common.legend = TRUE, legend = "bottom"))
```

### Interactive Plots
```{r include=FALSE}

# plotly::ggplotly(all_v_MHC)
```

```{r include=FALSE}
# plotly::ggplotly(diff_v_MHC)
```

# Session Information
## System Information
```{r}
sessioninfo::platform_info()
```


## Package Versions
```{r}
sessioninfo::package_info() %>% 
    filter(attached = TRUE) %>% 
    select(package, loadedversion, date, source) %>%
    knitr::kable(booktabs = TRUE, row.names = FALSE)
```