---
title: "Seq Data Characteristics"
date: '`r Sys.Date()`'
output: 
    bookdown::html_document2:
        toc: true
        toc_float: true
        df_print: paged
        code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(fastqcr)
```

# Background
Generate summary figures and tables with sequence data characteristics for fastq datasets provided to the challenge participants.

Fastq app on precisionFDA was used to calculate seq data summary statistics and the `fastqcr` R package was used to summarize the qc results.

Seqkit fq2tsv used to extract read quality and length information for PacBio and ONT data

# Fastqc results
## Loading Data
```{r}
qc_lst <- here("data","fastqc") %>% 
    list.files(pattern = "stats-fastqc.txt") %>% 
    set_names(str_remove(., pattern = ".stats-fastqc.txt")) %>% 
    map(~here("data","fastqc", .))

qc_metadata_df <- tibble(sample = names(qc_lst), fastqc_file = unlist(qc_lst)) %>% 
    mutate(Technology = case_when(str_detect(sample, "PacBio") ~ "PACBIO",
                                  str_detect(sample, "novaseq") ~ "ILLUMINA",
                                  str_detect(sample, "guppy") ~ "ONT",
                                  TRUE ~ "ERROR"),
           giab_id    = str_sub(sample, start = 0, end = 5),
           read_dir   = case_when(Technology %in% c("PACBIO", "ONT") ~ "R0",
                                  str_detect(sample, "R1") ~ "R1",
                                  str_detect(sample, "R2") ~ "R2"))

ill_metadata_df <- qc_metadata_df %>% 
    filter(Technology == "ILLUMINA")

pb_metadata_df <- qc_metadata_df %>% 
    filter(Technology == "PACBIO")

ont_metadata_df <- qc_metadata_df %>% 
    filter(Technology == "ONT")

modules_to_read <- c(
    "Basic Statistics",
    "Per base sequence quality",
    "Per sequence quality scores",
    "Sequence Length Distribution"
)

qc_ill <- qc_read_collection(ill_metadata_df$fastqc_file, 
                             sample_names = ill_metadata_df$sample,
                             modules = modules_to_read)
qc_pb <- qc_read_collection(pb_metadata_df$fastqc_file, 
                             sample_names = pb_metadata_df$sample,
                             modules = modules_to_read)

qc_ont <- qc_read_collection(ont_metadata_df$fastqc_file, 
                             sample_names = ont_metadata_df$sample,
                             modules = modules_to_read)
```

# Generating Summary Statistics Table

```{r}
bind_rows(qc_ill$basic_statistics, 
          qc_pb$basic_statistics) %>% 
    left_join(qc_metadata_df) %>% 
    filter(Measure %in% c("Total Sequences", "Sequence length", "%GC")) %>% 
    select(Technology, giab_id, Measure, Value, read_dir) %>% 
    pivot_wider(names_from = "Measure", values_from = "Value") %>% 
    ## Only using R1 for Illumina since R1 and R2 are the same
    filter(read_dir != "R2") %>% 
    select(-read_dir)
```

# Seq Quality Comparison
```{r}
qc_ill$per_sequence_quality_scores %>% 
    left_join(qc_metadata_df)  %>% 
    ggplot() + 
    geom_bar(aes(x = Quality, y = Count, fill = read_dir), 
             position = "dodge",
             stat = "identity") + 
    facet_wrap(~giab_id, ncol = 1) + 
    theme_bw() + labs(fill = "Read Direction", x = "Sequence Quality Score") + 
    theme(legend.position = "bottom")
```

## PacBio Plot

```{r}
pb_qual <- qc_pb$per_sequence_quality_scores %>% 
    left_join(qc_metadata_df)  %>% 
    ggplot() + 
    geom_bar(aes(x = Quality, y = Count),
             stat = "identity") + 
    facet_wrap(~giab_id, ncol = 1) + 
    theme_bw() + labs(x = "Sequence Quality Score") + 
    theme(legend.position = "bottom")

pb_length <- qc_pb$sequence_length_distribution %>% 
    left_join(qc_metadata_df)  %>%
    mutate(Length = as.integer(str_remove(Length, "-.*"))) %>% 
    filter(Length > 10000, Length < 20000) %>%
    ggplot() + 
    geom_bar(aes(x = Length/1000, y = Count),
             stat = "identity") + 
    facet_wrap(~giab_id, ncol = 1) + 
    theme_bw() + labs(x = "Read Length (kb)") + 
    theme(legend.position = "bottom")

(pb_combined_plt <- ggpubr::ggarrange(pb_qual, pb_length, labels = "AUTO"))
```
```{r}
ggsave(here("figures","pb_seq.pdf"), pb_combined_plt, width = 5, height = 5)
ggsave(here("figures","pb_seq.png"), pb_combined_plt, 
       width = 5, height = 5, dpi = "retina")
```

# Seqkit
## Load Data
```{r}
library(data.table)
library(dtplyr)
library(dplyr, warn.conflicts = FALSE)
```

```{r}
calc_n50 <- function(seq_lengths){
    sorted_lengths <- sort(seq_lengths)
    cum_lengths <- cumsum(as.numeric(sorted_lengths))
    min(sorted_lengths[cum_lengths >= max(cum_lengths)*0.5])
}

get_seq_stats <- function(seq_file){
    seq_df <- fread(seq_file, col.names = c("read_id","length", "qual"))
    
    ## Calculating Stats
    n_reads <- nrow(seq_df)
    n50 <- calc_n50(seq_df$length)
    med_length = median(seq_df$length)
    
    ## Return DF
    tibble(n_reads, n50, med_length)
}

seq_stats_df <- list.files(here("data", "fq_stats"), 
                           full.names = TRUE) %>% 
    set_names(basename(.)) %>% 
    map_dfr(get_seq_stats, .id = "dataset")
```

```{r}
seq_stats_df
```

# ONT Read Length  Plot
```{r}
get_ont_seq <- function(seq_file, subset_frac = 0.05, short_read_filter = 100){
    seq_dt <- fread(seq_file, col.names = c("read_id","length", "qual"))
    ## Excluding short reads
    seq_dt <- seq_dt[seq_dt$length > short_read_filter,]
    
    ## Random subset%
    n_seqs <- nrow(seq_dt)
    seq_subset <- sample(1:n_seqs, size = floor(n_seqs * subset_frac))
     
    seq_dt[seq_subset,] %>% 
        as_tibble() 
}

set.seed(100)
ont_fq_df <- list.files(here("data", "fq_stats"), 
                        pattern = "Guppy",
                        full.names = TRUE) %>% 
    set_names(basename(.)) %>% 
    map_dfr(get_ont_seq, .id = "dataset")

(
    read_len_fig <- ont_fq_df %>%

        mutate(GIAB = str_remove(dataset, "_.*")) %>%
        ggplot() +
        ggridges::geom_density_ridges(aes(
            y = GIAB,
            x = length
        )) +
        scale_x_log10() +
        theme_bw() +
        annotation_logticks(sides = "b") +
        labs(x = "Read Length (bp)",
             y = "GIAB ID") +
        theme_bw() + 
        theme()
)
```
```{r}
ggsave(here("figures","ont_seq_length.png"), read_len_fig, 
       width = 4, height = 4, dpi = "retina")
```


# Coverage analysis
```{r}
get_cov_stats <- function(cov_file){
    cov_df <- fread(cov_file, 
                    col.names = c("chrom","start","end","coverage")) %>% 
        as_tibble() %>% 
        filter(chrom %in% paste0("chr", 1:22)) 
    
    ## Calculating Stats
    coverage <- cov_df$coverage
    median_cov <- median(coverage)
    lq_cov <- quantile(coverage, probs = 0.25)
    uq_cov <- quantile(coverage, probs = 0.75)
    max_cov <- max(coverage)
    min_cov <- min(coverage)
    
    ## Return DF
    tibble(median_cov, lq_cov, uq_cov, max_cov, min_cov)
}

cov_stats_df <- list.files(here("data", "coverage"), 
                           full.names = TRUE) %>% 
    set_names(basename(.)) %>% 
    map_dfr(get_cov_stats, .id = "dataset")
```

```{r}
cov_stats_df
```


# Session Information
## System Information
```{r}
sessioninfo::platform_info()
```


## Package Versions
```{r}
sessioninfo::package_info() %>% 
    filter(attached = TRUE) %>% 
    select(package, loadedversion, date, source) %>%
    knitr::kable(booktabs = TRUE, row.names = FALSE)
```